\hyphenation{ALTA} % to avoid breaking ALTA over two lines

Welcome to the \textbf{23rd Annual Workshop of the Australasian Language Technology Association (ALTA 2025)}. Hosted on the campus of the University of Sydney, ALTA 2025 will provide a platform for the exchange of ideas, exploration of innovations, and discussion of the latest advancements in language technology. The workshop acknowledges the significance of its location on the traditional lands of the Gadigal people of the Eora nation, underscoring a commitment to inclusivity and respect. ALTA 2025 convenes leading researchers, industry experts, and practitioners in the fields of natural language processing (NLP) and computational linguistics.\newline

The program at ALTA this year reflects a wide variety of topics of relevance to contemporary NLP research.  This year, ALTA had three options for submission: direct archival papers (long and short), direct non-archival papers (various lengths), and a commitment site for papers reviewed through ACL Rolling Review (ARR). We received 24 archival submissions, accepting 13 after double-anonymous reviews by 3 reviewers per paper. This gives an overall acceptance rate of 54\%, which is comparable to previous iterations of ALTA. The papers will be presented in four thematic sessions: natural language understanding, natural language generation, robustness and security of LLMs, and multimodality.\newline

This year, we introduced an alternative to present papers published at international conferences such as ACL, EMNLP, with the view to enable ALTA to be a local forum to disseminate internationally acknowledged research and facilitate discussions among researchers in the region. This was in addition to non-archival papers which may be early work or research proposals. We received 14 submissions in the non-archival track and accepted 13 of them. This includes 8 papers that have been published elsewhere. Every non-archival submission was reviewed by one reviewer. The non-archival papers will be presented as lightning talks across three thematic sessions.\newline

The workshop includes a tutorial on ‘Alignment of Large Language Models with Human Preferences and Values’ by presenters from Macquarie University and a shared task on linking adverse drug events to medical ontologies. This year’s panel discussion is on ‘Teaching/Using NLP for teaching’ and will be moderated by Aditya Joshi. We thank our panelists for offering their insights on the topic. We thank the tutorial presenters, shared task chair and participants. The shared task participants will also present short talks at the workshop.\newline

This year’s submissions showcase advancements across a wide array of topics. The thematic sessions highlight the range of topics covered in the workshop. Scientific literature, code and social media data are recurring text forms. The papers on language and multimodality highlight the narrowing gap between different modalities as foundation models progress towards unified modeling.\newline

We would like to extend our heartfelt gratitude to our keynote speakers: Prof. Alistair Knott - (Victoria University of Wellington), Prof. Mark Johnson (Macquarie University and Oracle). A special thank you to Prof. Massimo Piccardi (University of Technology Sydney) who was the judge for the best paper awards.\newline

Thank you to the members of the organising committee and volunteers for their hard work in preparing and running ALTA. We extend our heartfelt appreciation to the reviewers for their effort in writing detailed and constructive reviews.\newline

Lastly, ALTA 2025 gratefully acknowledges the support of our sponsors: Google (Gold tier), Amazon Nova and Mongo DB (Silver tier), and GE Aerospace and University of Melbourne (Bronze tier). We are also proud to have the University of Sydney as our host.\newline

The success of this workshop would not be possible without your invaluable contributions. Welcome to USyd and Sydney! We hope that you enjoy ALTA 2025, and look forward to a rewarding and inspiring time together.\newline

\hspace*{\fill}Jonathan K. Kummerfeld\newline
\hspace*{\fill}\textit{ALTA 2025 General Chair}\newline\newline
\hspace*{\fill}Aditya Joshi\newline
\hspace*{\fill}Mark Dras\newline
\hspace*{\fill}\textit{ALTA 2025 Program Chairs}